---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Step 3: Create two visualizations of your data set that illustrate some hidden information or insight into the data set. Make sure that the visualizations are easy for your viewers to understand and follow best practices. Do this by Oct. 20, 2023 at the start of class. You will submit a compressed folder containing your Rmd file and data set to Canvas by the deadline.
#Loading in the data
```{r}
getwd()
#setwd("C:/Users/mirab/Desktop/CPSC292")
getwd()
tsdata <- read.csv("taylor_swift_spotify.csv")
```

```{r}
head(tsdata)
```

#raw danceability data
```{r}
library(ggplot2)
# Create a scatter plot
ggplot(tsdata, aes(x = album, y = danceability)) +
  geom_point() +  # Scatter plot
  xlab("Album") + ylab("Danceability") +  # Labels for axes
  theme_minimal()+  # Use a minimal theme 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#updated danceability data (colorblind friendly)
```{r}
library(ggplot2)

# Define a vector of albums to include
included_albums <- c("Taylor Swift", "Fearless (Taylor's Version)", "Speak Now (Taylor's Version)", "Red (Taylor's Version)", "reputation", "Lover", "1989", "folklore", "evermore", "Midnights")

# Filter the data to include only the specified albums
tsdata_filtered <- tsdata[tsdata$album %in% included_albums, ]

# Define colors for each album
album_colors <- c(
  "Taylor Swift" = "#44AA99",
  "Fearless (Taylor's Version)" = "#DDCC77",
  "Speak Now (Taylor's Version)" = "#332288",
  "Red (Taylor's Version)" = "#882255",
  "reputation" = "black",
  "Lover" = "#CC6677",
  "1989" = "#88CCEE",
  "folklore" = "grey",
  "evermore" = "#117733",
  "Midnights" = "#AA4499")

# Create a scatter plot with the filtered data and custom colors
plot <- ggplot(tsdata_filtered, aes(x = album, y = danceability, color = album)) +
  geom_point(size = 4) +  # Increase point size
  scale_color_manual(values = album_colors) +  # Assign custom colors
  xlab("Album") + ylab("Danceability")+ ggtitle("Danceability of Each Taylor Swift Album")+ # Labels for axes
  theme_minimal() +  (theme(plot.title = element_text(hjust = 0.5))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate and space out x-axis labels
  theme(legend.position = "none")  # Remove the legend

# Print the plot
print(plot)

```

#boxplot
```{r}
# Load the ggplot2 library
library(ggplot2)

# Define a vector of albums to include
included_albums <- c("Taylor Swift", "Fearless (Taylor's Version)", "Speak Now (Taylor's Version)", "Red (Taylor's Version)", "reputation", "Lover", "1989", "folklore", "evermore", "Midnights")

# Filter the data to include only the specified albums
tsdata_filtered <- tsdata[tsdata$album %in% included_albums, ]

# Define colors for each album
album_colors <- c(
  "Taylor Swift" = "#44AA99",
  "Fearless (Taylor's Version)" = "#DDCC77",
  "Speak Now (Taylor's Version)" = "#332288",
  "Red (Taylor's Version)" = "#882255",
  "reputation" = "black",
  "Lover" = "#CC6677",
  "1989" = "#88CCEE",
  "folklore" = "grey",
  "evermore" = "#117733",
  "Midnights" = "#AA4499")

# Create a boxplot with the filtered data and custom colors
plot <- ggplot(tsdata_filtered, aes(x = album, y = danceability, fill = album)) +
  geom_boxplot() +  # Create a boxplot
  scale_fill_manual(values = album_colors) +  # Assign custom colors
  xlab("Album") + ylab("Danceability") + ggtitle("Danceability of Each Taylor Swift Album") +  # Labels for axes and title
  theme_minimal() +  (theme(plot.title = element_text(hjust = 0.5))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate and space out x-axis labels
  theme(legend.position = "none")  # Remove the legend

# Print the plot
print(plot)
```





```{r}
library(ggplot2)

# Define the album you want to analyze
target_album <- "Lover"

# Filter the data to include only the tracks from the target album
album_data <- tsdata[tsdata$album == target_album, ]

# Create a bar plot of danceability for each track (name)
ggplot(album_data, aes(x = name, y = danceability)) +
  geom_bar(stat = "identity", fill = "#CC6677", color = "black") +
  xlab("Track Name") + ylab("Danceability") +
  theme_minimal() + (theme(plot.title = element_text(hjust = 0.5)))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle(paste("Danceability of Tracks on", target_album)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, by = 0.1))
```
